All watched over by machines of loving grace

### 2021-06-22 paper/paper.tex
paper: more work on introduction

Shifting focus, we see an interesting analogy.
Typical commodity operating systems---such as Microsoft Windows and desktop Linux distributions---also implement a \emph{kernel}.
This kernel is also the sole system component that has unfettered access to the host machine's hardware\footnote{Barring other privileged systems software, such as the hypervisor, which we consider as another form of operating system}, and will typically provide various abstractions, such as a hierarchically-structured filesystem, sockets, and so on, to untrusted, user-space code.
System-wide security policies are enforced by the kernel using a variety of mechanisms, including password-protected user accounts, file and directory permissions, access control lists, firewalls, and runtime-monitoring facilities.
From a security point-of-view, an operating system kernel again acts as the \emph{root of all trust} within the system.

At heart---though both types of kernels are intended for very different use-cases---both types of kernel are intended to enforce a series of \emph{kernel invariants} even in the face of unbounded interaction with untrusted code.
However, despite this common cause, the implementation techniques employed by authors of the two type of kernels tends to be very different.
In particular, an LCF-style kernel is little other than a library implemented in a high-level programming language---typically some variant of ML, such as Standard ML or OCaml.
Invariants are maintained by the kernel judiciously exposing functionality as an API that offers functions to manipulate elements of a collection of abstract types, thereby exploiting the soundness of the implementation language's type-system.
Untrusted code then interacts with the kernel through this API.

@@

Supervisionary implements a version of Gordon's higher-order logic (or HOL, henceforth), a close relative of Church's Simple Theory of Types that extends Church's system with polymorphic types, albeit with no ML-style \texttt{let}-polymorphism.
We describe HOL here, introducing enough material for the reader to understand the remainder of the paper---motivating HOL or describing its expressiveness is outside of the scope of this paper.

>>>>

Shifting focus, we see an interesting analogy: desktop operating systems---such as Microsoft Windows and desktop Linux distributions---also implement a \emph{kernel}.
This kernel is also the sole system component that has unfettered access to the host machine's hardware\footnote{Barring other privileged systems software, such as the hypervisor, which we consider as another form of operating system}, and will typically provide various abstractions, such as a hierarchically-structured filesystem, sockets, and so on, to untrusted user-space code.
System-wide security policies are enforced by the kernel using a variety of mechanisms, including password-protected user accounts, file and directory permissions, access control lists, firewalls, and runtime-monitoring facilities.
From a security point-of-view, an operating system kernel again acts as the \emph{root of all trust} within the system.

At heart---though intended for very different use-cases---both types of kernel are intended to enforce a series of \emph{kernel invariants} even in the face of unbounded interaction with untrusted code.
However, despite this common cause between the two families of systems, the protection mechanisms used to enforce these invariants tend to be very different.

In particular, an LCF-style kernel is little other than a library implemented in a high-level programming language---typically some variant of ML, such as Standard ML or OCaml.
Invariants are maintained by the kernel judiciously exposing functionality through an API---offering functions to manipulate elements of certain abstract types---and through which untrusted code interacts with the kernel.
Logical soundness is therefore contingent on the type soundness of the implementation meta-language's type-system.

In contrast, desktop operating systems use hardware separation mechanisms to protect various \emph{kernel objects}---that is, book-keeping structures used by the kernel---from unfettered access by untrusted code.
These hardware separation mechanisms include elevated \emph{privilege rings} (or \emph{exception levels})---granting the kernel access to protected hardware resources, such as system registers---and memory protection mechanisms through which the kernel can ``carve out'' a private scratch space to safely stash kernel objects.

@@

Supervisionary implements a version of Gordon's higher-order logic (or HOL, henceforth), a close relative of Church's Simple Theory of Types that extends Church's system with top-level polymorphism.
We describe HOL here, introducing enough material for the reader to understand the remainder of the paper---motivating HOL or describing its expressiveness is outside of the scope of this paper.
